Diverse snow algae blooms in the alpine and subalpine zones of British Columbia

### 2020-01-08 manuscript/manuscript.Rmd
final to ky

---
title: "Diverse snow algae blooms in the alpine and subalpine zones of British Columbia"
author:
  - name: Casey B. Engstrom
    institute: SFU
  - name: Kurt M. Yakimovich
    institute: SFU
  - name: Lynne M. Quarmby
    institute: SFU
    email: quarmby@sfu.ca
institute:
  - SFU: Dept. Molecular Biology and Biochemistry, Simon Fraser University, 8888 University Drive, Burnaby, BC, V5A 1S6, Canada
output: 
  bookdown::word_document2:
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=templates/scholarly-metadata.lua'
      - '--lua-filter=templates/author-info-blocks.lua'
    reference_docx: templates/word-template.docx
  # pdf_document: 
  #   fig_caption: yes
  #   number_sections: yes
csl: frontiers.csl
bibliography: refs.bib
abstract: |
  Snow algae blooms cover vast areas of summer snowfields worldwide, and can reduce albedo and increase snow melt. Despite their global prevalence, little is known about the species that comprise snow algae blooms. We used 18S and *rbcL* high-throughput amplicon sequencing (metabarcoding) to survey 33 snow algae samples from alpine and subalpine habitats in the Coast Range of British Columbia, Canada. Species composition varied from bloom to bloom: *Sanguina* predominated above treeline, while *Chloromonas* were prevalent at lower elevations. *Chlainomonas* were abundant in samples across all elevations. The highest diversity was contained within the genus *Chloromonas*, which included operational taxonomic units (OTUs) that likely represent novel species of snow algae. ???

@@

# Introduction {-}
<!-- global, vast blooms. albedo->snowmelt. what species? -->
Each summer in polar and alpine snowfields worldwide, vast areas of snow surface are colored red by snow algae blooms. Such blooms are reported from every continent [@segawa_bipolar_2018;@duval_snow_1999;@marchant_snow_1982;@yoshimura_community_1997;@vimercati_nieves_2019] and Arctic sea ice [@gradinger_snow_1996]. They can be extensive: in Alaska, snow algae spectral signatures were detected in one third of a 1,900 km^2^ icefield [@ganey_role_2017]. In recent years snow algae have received attention for their role in reducing snow surface albedo, which could substantially impact snow melt rates [@lutz_biogeography_2016;@ganey_role_2017]. Thus, snow algae could impact the timing of spring melt and runoff, reducing water supplies in mountain snowpacks and glacier mass balance. Despite their global ubiquity, we are only beginning to identify the species that cause snow algae blooms.

@@

l

@@

<!-- formatting indentation? -->

@@

**Figure 2.** Ordination plots of *rbcL* amplicon sequence variant (ASV) similarity. A. Multidimensional scaling (MDS) plot. Taxonomy (assigned by IDTaxa) indicated by color, and size of point is proportional to cumulative relative abundance. Dotted ellipses show OTU clusters. Stress = 0.13. B. t-Distributed Stochastic Neighbor Embedding (t-SNE) dimensionality reduction plot of snow algae *rbcL* ASVs. Perplexity = 30.

>>>>

---
title: "Diverse snow algae blooms in the alpine and subalpine zones of British Columbia"
author:
  - name: Casey B. Engstrom
    institute: SFU
  - name: Kurt M. Yakimovich
    institute: SFU
  - name: Lynne M. Quarmby
    institute: SFU
    email: quarmby@sfu.ca
institute:
  - SFU: Dept. Molecular Biology and Biochemistry, Simon Fraser University, 8888 University Drive, Burnaby, BC, V5A 1S6, Canada
output: 
  bookdown::word_document2:
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=templates/scholarly-metadata.lua'
      - '--lua-filter=templates/author-info-blocks.lua'
    reference_docx: templates/word-template.docx
  # pdf_document: 
  #   fig_caption: yes
  #   number_sections: yes
csl: frontiers.csl
bibliography: refs.bib
abstract: |
  Snow algae blooms cover vast areas of summer snowfields worldwide, and can reduce albedo and increase snow melt. Despite their global prevalence, little is known about the species that comprise snow algae blooms. We used 18S and *rbcL* high-throughput amplicon sequencing (metabarcoding) to survey 33 snow algae samples from alpine and subalpine habitats in the Coast Range of British Columbia, Canada. Species composition varied from bloom to bloom: *Sanguina* predominated above treeline, while *Chloromonas* were prevalent at lower elevations. *Chlainomonas* were abundant in samples across all elevations. The highest diversity was contained within the genus *Chloromonas*, which included operational taxonomic units (OTUs) that likely represent novel species of snow algae.

@@

# Introduction {-}
<!-- global, vast blooms. albedo->snowmelt. what species? -->
Each summer in polar and alpine snowfields worldwide, vast areas of snow surface are colored red by snow algae blooms. Such blooms are reported from every continent [@segawa_bipolar_2018;@duval_snow_1999;@marchant_snow_1982;@yoshimura_community_1997;@vimercati_nieves_2019] and Arctic sea ice [@gradinger_snow_1996]. They can be extensive: in Alaska, snow algae spectral signatures were detected in one third of a 1,900 km^2^ icefield [@ganey_role_2017]. In recent years snow algae have received attention for their role in reducing snow surface albedo, which could substantially impact snow melt rates [@lutz_biogeography_2016;@ganey_role_2017]. Thus, snow algae could impact the timing of spring melt and runoff, depleting summer water supplies held in mountain snowpacks and reducing glacier mass balance. Despite their global ubiquity, we are only beginning to identify the species that cause snow algae blooms.

@@

@@

<!-- formatting indentation??? -->

@@

**Figure 2.** Ordination plots of *rbcL* amplicon sequence variant (ASV) similarity. A. Multidimensional scaling (MDS) plot. Taxonomy (assigned by IDTaxa) indicated by color, and size of point is proportional to cumulative relative abundance. Dotted ellipses show OTU clusters. Stress = 0.13. B. t-Distributed Stochastic Neighbor Embedding (t-SNE) dimensionality reduction plot of snow algae *rbcL* ASVs. Perplexity = 30.l
